---
title: "Applied Data Science Project"
author: "Alex Hyman"
date: "8/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading in the data

Reading in the data frame and converting the customer ID into the string instead of a factor. Also loading in ggplot for future plots.

```{r cars}
df <- read.csv("data.csv")
df$customerID <- as.character(df$customerID)
library(ggplot2)
```

## Churn Rates

Is there a difference in churn rates between all customers and the churn rates in senior citizens?

```{r}
#Churn rates among senior citizens
seniors <- length(df$customerID[(df$SeniorCitizen == 1 & df$Churn == "Yes")]) / length(df$customerID[df$SeniorCitizen == 1]) * 100
#Rounding the rate
seniors = round(seniors, digits = 2)
#Print out churn rate for senior citizens
cat("Churn rate among senior citizens: ", seniors, "%", sep = "")

#Churn rate for non-senior citizen customers
allOtherCustomers <- length(df$customerID[df$Churn == "Yes" & df$SeniorCitizen == 0]) / length(df$customerID[df$SeniorCitizen == 0]) * 100
allOtherCustomers <- round(allOtherCustomers, 2)
cat("Churn rate of non-Seniors: ", allOtherCustomers, "%", sep = "")

#Creating booleans for churn in t.test application
df[df$Churn == "Yes", "LogicalChurn"] <- T
df[df$Churn == "No", "LogicalChurn"] <- F
#Are these groups statistically different
t <- t.test(LogicalChurn ~ SeniorCitizen, data = df)
print(t)
```

The churn rates for senior citizens are statistically greater than those that are not senior citizens.

## Visualizing Churn by Average Cost and Tenure

Can a pattern in churn be recognized based on how much is spent and what kind of contract? 

```{r}
#Plotting tenure vs charge to see if pattern in where churn is
g <- ggplot(df, aes(tenure, MonthlyCharges, color = Churn, shape = Contract))
g <- g + geom_point()
g <- g + ggtitle("Tenure vs. Average Monthly Charges")
g <- g + xlab("Tenure") + ylab("Average Monthly Charges")
g

p <- ggplot(df[df$Contract == "Month-to-month", ], aes(tenure, fill = Churn))
p <- p + geom_histogram(binwidth = 1) + xlab("Tenure (months)")
p <- p + ggtitle("Histogram on Churn and tenure\nfor Month-to-Month Customers")
p
```
Appears to be separated in a diagonal manner.

## Test for independence in contract

Is there any relationship between the cnatract type and churn rate?

```{r}
#Chi-square test for independence
chit <- chisq.test(df$Contract, df$Churn)
#What was actually observed
chit$observed
#What was expected
chit$expected
```
There is definitely some relationship between the contract type and churn rate. Two year contract is significantly below the expected rate.

```{r}
mean(df[df$Churn == "Yes", "tenure"] )
median(df[df$Churn == "Yes", "tenure"] )
```

```{r}
p <- ggplot(df, aes(Contract)) + geom_bar(aes(fill = Churn))
p <- p + ylab("Count") + ggtitle("Churn by Contract type")
p
```




```{r}
tapply(df$customerID, df$PhoneService == "Yes" & df$Churn == "Yes", length)
```

